---
title: "ESM 206 Lab 5"
author: "Peyton Moore"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Lab 5 objectives 

- Learn to parse dates with 'lubridate' 
- Get counts of obsercations with count()
- Uncount 
- One and two-sample t-test 
- Create a heat map with geom_tile()

```{r, include = FALSE}
# Read in data and attach packages 
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data","lobster_abundance.csv"), 
                              na = "-99999") %>% 
  janitor::clean_names()
  

```


Use tidyr::uncount() function to convert our lobster data from frequency format to case format 

```{r}
lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)

```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site), 
              alpha = 0.5, 
              width = 0.2)


# Histogram

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")
  
ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)


```


#### Convert the 'date' column to class 'date'

We'll use the lubridate package to convert to Date format and then to help us easily parse month and year 

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

#class(lobster_date$date)
```


Now lets parse year and month using lubridate::month() and lubridate::year():

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
   obs_month = lubridate::month(date_new, label = TRUE), 
   obs_year = lubridate::year(date_new)
    
  )
```

Now lets find counts of observed lobsters based on different hierarchical groupings: 


first lets count lobsters by year and month
```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_y <-lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_site <-  lobster_parse_date %>% 
  dplyr::count(site)


```


If we wwant to create a summary table that contains statistics OTHEr than counts by group, its easier to use group_by() + n()

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarise(
    mean_size = mean(size_mm, na.rm = TRUE), 
    sd_size = sd(size_mm, na.rm = TRUE), 
    lobster_number = n()
    
  )

```

#### Find confidence intervals 

Use t.test function to find confidence intervals for one sample and perform t tests to compare means of two samples


```{r}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)


t.test(ivee_lobsters)



```


#### Two sample t test to compare means 

We want to ask, is there a signifcant difference in lobster lengths at Naples and Mohawk reefs?

We've sone our necessary and exploratory analyses to determine that a 2 sampls t test for means comparison is appropriate. 

```{r}
napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)


mn_ttest <- t.test(mohk_sample, napl_sample)
mn_ttest



```

There is a significant difference between lobster lengths between Naples and Mohawk Reef. (t(df) = statistic, p <0.001, alpha = 0.05).

ˆ
```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL","MOHK"))


mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic
```


Here is the p-value: `r mn_ttest2$p.value`

Here is the t-statistic: `r mn_ttest2$statistic`




#### Now a heatmap!

```{r}
lobster_ys <-  lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```





